Assessing reliability of an online behavioural laterality battery: a pre-registered

### 2020-07-03 draft_notebook_writeup.Rmd
Drafted majority of results. Still requires minimal trials for stable LI which is next to do.

Department of Experimental Psychology, University of Oxford, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford OX2 6GG

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

@@

Prior to data collection, we estimated that we would need 300 participants in order to achieve sufficient precision of the estimate of reliability. With this sample size, if the true test-retest correlation is .58, then the 95% confidence interval around the estimate will be .50-.65.  Thus, with this sample size, if we observed a correlation of .65, then the true correlation is unlikely to be lower than .58. As will become clear in the Data Analysis section, our final sample exceeded 300 participants. In total 441 native English speakers were initially recruited through Prolific Academic. Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g. head injury, brain tumour, stroke) or significant hearing loss/impairment.

@@

@@

The basic demographics questionnaire asked participants to report handedness, footedness, years in education, and whether they were bilingual.

@@

The Edinburgh Handedness Inventory (Oldfield, 1971) was administered to quantify handedness on a continuum. For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred). Scores range between -100 and 100 with negative values indicating a left-hand preference and positive values a right hand preference.

@@

The Test for Advanced Learners of English (Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge. Participants saw 60 letter strings and had to indicate which word they knew. Of the 60 strings, 40 were actual English words and 20 were non-words. LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types. Accordingly, the LexTALE scores ranged from 0 to 100.

@@

```{r read_data,echo=FALSE,include=TRUE}
# read in data from csv
LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
# cor test for rhyme 
corR<- cor(LIs$Day2.RDT_RT_LI, LIs$Day1.RDT_RT_LI, use = "complete.obs")
# table of counts for footed and sigh
short_dat <- dplyr::select(LIs, "subject", "handedness", "footedness", "miles", "porta")
long_dat <- gather(short_dat, task, response, footedness:porta, factor_key=TRUE)
# make table
tab <- as.data.frame(table(long_dat$handedness, long_dat$task, long_dat$response))
tab <- tab %>% 
  rename(
    Handedness = Var1,
    Task = Var2,
    Response= Var3,
    Count= Freq
    )
split <- ggplot(data=tab, aes(x=Response, y=Count, fill=Handedness)) +
  geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~Task) +
    geom_text(aes(label=Count), vjust=-0.3, position = position_dodge(0.9), size=3.5)+
  theme_classic() + theme(legend.position = "top") + ylab("Count") + xlab("") + ylim(0,220)
ggsave(
  "handed_sight.png",
  plot = split,
  width = 6, height = 4,
  dpi = 300
)
```

Out of the 441 partiipants recuited via Prolific, 421 (`r round(421/441*100, 1)`%) complete both sessions. We then removed 21 (`r round(21/421*100, 1)`%) participants from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency. A further 10 participants (`r round(10/400*100, 1)`%) were excluded due to issues with long refresh rates or self report that stimuli were displaying on several tasks. This left a sample of 390 participants.

For each task we adaopted the following data cleaning procedures. Applying each of these meant that the number of participants left in the final analyses could vary by task. First, for all tasks, we removed trials in which reaction times were shorter than 200 ms. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as reaction times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)). Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for reaction time, since reaction time is always non-negative, and so most outliers would be in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participants’ data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

Next, for the chimeric faces task, we removed partcipants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemisface. These data trimming procedures left data for `r 390-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces. For the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear. These data trimming procedures left data for `r 390-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.

With regards to the rhyme task, we removed participants who scored less than 50% accuracy in either visual field. Furthermore, we removed particpants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme). These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on the rhyme decision task. Seeing as we were removing a relatively high portion of participants, want wanted to check whether this task was yielding reliably results. With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(corR,3)`, which is far smaller than the r= 0.65 that we pre-registered as our criteria for a reliable task. Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme, we chose to end data collection here and analyses the 390 participants that we had collected under the assumption that the rhyme decision is not reliable. The final group of 390 participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female). `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual. `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual. The division of footedness and ocular dominance in left- and right-handers are shown in Figure 3. Below we describe the analysis procedures used to examine our research questions and hypotheses.

@@

For the rhyme task, we pre-registered that we would calculate LIs based on reaction time data, as accuracy during piloting was at ceiling. We specified that we would calculate LIs in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be calculated in a way that corresponded to a *t*-value. This was achieved by conducting a *t*-test for each participant where accuracy response times are the dependent variable and visual field is the independent variable. This would estimate the sensitivity to stimuli presented in either visual field and equate to a laterality index. One benefit of this approach is that we could identify participants who show a statistically significant reaction time advantage for one side, i.e. where the *t*-value on a 2-sided test has *p*< .05.

@@

For the reliability assessment, we first plotted the distributions of LIs and checked them for normality using a Shapiro-Wilk test. As the LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.

@@

1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1
1

>>>>

Department of Experimental Psychology, University of Oxford, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road, Oxford, OX2 6GG.

<!--- Leaving blank for now as not sure how corresponding authorship works in OSCCI lab --->
*Correspondence regarding this article should be addressed to XXXXXX, Department of Experimental Psychology, University of Oxford, Anna Watts Building, Woodstock Road, Oxford, OX2 6GG. Email: XXXXXX

@@

As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least .65. To achieve sufficient precision of the estimate of reliability, we estimated we needed 300 participants. With this sample size, if the true correlation is .58, then the 95% confidence interval around the estimate will be .50-.65. Thus with this sample size, if we observe a correlation of .65, then the true correlation is unlikely to be lower than .58.

Out of the 300 participants, we originally specified that we'd collected data for least 80 left-handers. We had examined the power that a sample of 80 participants would provide to make the originally planned model comparisons outlined under hypothesis 3. By comparing model fit using a likelihood test, 80 participants would provide 100% power for model (a) and 97% power for model (b). Thus, we would have sufficient power to assess the fit of each model using Akaike weights, which are more conservative than likelihood tests.

Despite setting out to recruit 300 participants, our final sample exceeded this target (we say more about this below). In total, 441 native English speakers were initially recruited through Prolific Academic. Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g. head injury, brain tumour, stroke) or significant hearing loss/impairment.

@@

<!--Dorothy, just a note that all Ps came from Prolific in the end-->

@@

The basic demographics questionnaire asked participants to report handedness, years in education, and whether they were bilingual. We also asked participnats a question about footedness: which foot do you normally use to step up on a ladder/step? Participants could answer "left", "right", or "don't know".

@@

The Edinburgh Handedness Inventory (Oldfield, 1971) was administered to quantify handedness on a continuum. For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred). The Edinburgh handedness inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference across ten tasks. For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed. For analysis of extreme left-handers, we focussed on left-handers with an index of -90 or less, which corresponds to near-exclusive use of the left hand.

@@

The Test for Advanced Learners of English (Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge. Participants saw 60 letter strings and had to indicate which word they knew. Of the 60 strings, 40 were actual English words and 20 were non-words. LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2). Accordingly, the LexTALE scores ranged from 0 to 100.

@@

```{r read_dat, echo- FALSE, include= FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)

# read in data from csv
LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
```

```{r prepdat,echo=FALSE,include=FALSE}
# remove those without day 2
# reformat NA
LIs$day2.complete <- ifelse(is.na(LIs$day2.complete), 0, LIs$day2.complete)
# remove
LIs_c <- LIs[LIs$day2.complete == 1,]
LIs <- LIs[LIs$day2.complete == 1,]
# lextale out df
LIs_no <- LIs[LIs$lexOut != "keep",]
# now remove them from LI
LIs <- LIs[LIs$lexOut == "keep",]
# table of counts for footed and sigh
short_dat <- dplyr::select(LIs, "subject", "handedness", "footedness", "miles", "porta")
long_dat <- gather(short_dat, task, response, footedness:porta, factor_key=TRUE)
# make table
tab <- as.data.frame(table(long_dat$handedness, long_dat$task, long_dat$response))
tab <- tab %>% 
  rename(
    Handedness = Var1,
    Task = Var2,
    Response= Var3,
    Count= Freq
    )
split <- ggplot(data=tab, aes(x=Response, y=Count, fill=Handedness)) +
  geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~Task) +
    geom_text(aes(label=Count), vjust=-0.3, position = position_dodge(0.9), size=3.5)+
  theme_classic() + theme(legend.position = "top") + ylab("Count") + xlab("") + ylim(0,220)
ggsave(
  "handed_sight.png",
  plot = split,
  width = 6, height = 4,
  dpi = 300
)
#Add a category of handedness that distinguishes extreme left from other left-handers
#In preregistration, extreme left defined as EHI laterality -90 or less.
LIs$handed3 <- LIs$handedness
w<-which(LIs$index_EHI<(-89))
LIs$handed3[w]<-'Extreme.left'
```

Out of the 441 partiipants recuited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) complete both sessions. We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency. This left a final sample of `r nrow(LIs)` participants.

For each task we adaopted the following data cleaning procedures. Applying each of these procedures meant that the number of participants left in the final analyses could vary by task. First, for all tasks, we removed trials in which reaction times were shorter than 200 ms. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as reaction times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)). Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for reaction time, since reaction time is always non-negative, and so most outliers would be in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participants’ data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

Next, for the chimeric faces task, we removed partcipants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemisface. These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces. For the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear. These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.

With regards to the rhyme task, we removed participants who scored less than 50% accuracy in either visual field. Furthermore, we removed particpants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme). These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on the rhyme decision task. Seeing as we were removing a relatively high portion of participants, want wanted to check whether this task was yielding reliably results. With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),3)`, which is smaller than the *r*= 0.65 that we pre-registered as our criteria for a reliable task. Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme decision task, we chose to end data collection here and analyse the `r nrow(LIs)` participants that we had collected under the assumption that the rhyme decision is not reliable. The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female). `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual. The division of footedness and ocular dominance in left- and right-handers are shown in Figure 3. Below we describe the analysis procedures used to examine our research questions and hypotheses.

@@

For the rhyme task, we pre-registered that we would calculate LIs based on reaction time data, as accuracy during piloting was at ceiling. We specified that we would calculate LIs in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be calculated in a way that corresponded to a *t*-value. This was achieved by conducting a *t*-test for each participant where accuracy response times are the dependent variable and visual field is the independent variable. This would estimate the sensitivity to stimuli presented in either visual field and equate to a laterality index. One benefit of this approach is that we could identify participants who show a statistically significant reaction time advantage for one side, i.e. where the *t*-value on a 2-sided test has *p*< .05. Based on considerations from pilot data, we planned to transform RT data from the RDT task by log-transformation before computing *t*-values.

@@

For the reliability assessment, we plotted the distributions of LIs and checked them for normality using a Shapiro-Wilk test. As the LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.

For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .75)^1^. This will allow us to identify the most efficient way to administer these tasks in future, while retaining good reliability. However, the full data (including all trials) was used for all subsequent analyses.

@@

```{r descriptives, echo=FALSE,include=FALSE}
# here we present the raw count and RT scores for each task
# read CF data
for_figs <- read.csv("./data/for_figs.csv", na.strings = "NA")
# get CF
CF.short.dat <- for_figs[for_figs$exclude== 0 | for_figs$exclude== 2,]
  CF.short.dat <-CF.short.dat[CF.short.dat$task== "CF",]
# read Dichotic data
DL.short.dat <- for_figs[for_figs$exclude== 0 | for_figs$exclude== 1,]
  DL.short.dat <- DL.short.dat[DL.short.dat$task== "DL",]
# read rhyme data
RD.short.dat <- for_figs[for_figs$exRDT== 0,]
  RD.short.dat <- RD.short.dat[RD.short.dat$task== "RD",]
# read finger data
finger.short.dat <- read.csv("./data/finger.csv", na.strings = "NA")

# plots
CF <- ggplot(CF.short.dat, aes(x = side, y = accurate)) +
        ggpirate::geom_pirate(aes(colour = side)) +
        theme_classic() + ylab("Count of correct responses") + xlab("") + facet_wrap(.~Session)+ 
        ggtitle("C") + theme(plot.title = element_text(hjust = 0.5))
DL <- ggplot(DL.short.dat, aes(x = side, y = accurate)) +
        ggpirate::geom_pirate(aes(colour = side)) +
        theme_classic() + ylab("Count of correct responses") + xlab("") + facet_wrap(.~Session)+ 
        ggtitle("B") + theme(plot.title = element_text(hjust = 0.5))
RD <- ggplot(RD.short.dat, aes(x = side, y = RT)) +
        ggpirate::geom_pirate(aes(colour = side)) +
        theme_classic() + ylab("Reaction time (ms)") + xlab("") + facet_wrap(.~Session) + 
        ggtitle("A") + theme(plot.title = element_text(hjust = 0.5))
FT <- ggplot(finger.short.dat, aes(x = side, y = sequences)) +
        ggpirate::geom_pirate(aes(colour = side)) +
        theme_classic() + ylab("Count of correct sequences") + xlab("") + facet_wrap(.~Session)+ 
        ggtitle("D") + theme(plot.title = element_text(hjust = 0.5))
all.plot <- suppressMessages(suppressWarnings(ggpubr::ggarrange(RD, DL, CF, FT)))
all.plot
# save
ggsave(
  "all.vf.png",
  plot = all.plot,
  width = 6, height = 6,
  dpi = 300
)
```

Participants' average reaction times on the rhyme decision task are shown in Figure 4 broken down by visual field and session. While we do not analyse accuracy, accuracy on the rhyme task was high  (*M*= `r round(mean(RD.short.dat$p.corr, na.rm= TRUE),1)`%, *SD*= `r round(sd(RD.short.dat$p.corr, na.rm= TRUE),2)`%). The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 4, along with the number of sequences gerated by the left and right hand. LIs were computed in several ways. For all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)). It is important to note, however, that we present LIs such that they are reported to be consistent with lateralisation at the brain level (i.e. by multiplying LIs by -1). A negative LI indicates left lateralisation at the brain level while a positive LI indicates right lateralisation at the brain level. For the rhyme decision task, we then computed an LI using a *t*-test for inidiviaul participants fit to log-tranformed reaction time data. For the reamining tasks (which made use of count data) we computed a *z*-statistic for each participant using count data.

![Figure 4. Mean scorees across the two sessions of the behavioural laterality tasks. (A) Reaction times in each visual field on the rhyme decision tasks. (B) Count of correctly identified sounds played to the left or right ear. (C) Count of correctly identified emotions in each hemiface. (D) Count of sequences generated by the left and right hand.](all.vf.png)

For each measure, LIs and their test-retest correlation coefficients are reported in Table 1. In addition, Table 1 inlcudes the *p*-values for a one-sampled test and Shapiro-Wlk test for each measure  As the distirbution of each LI was non-normal, spearman's rank correlations were used to estimate the test-retest reliability. Confidence intervals were estimated using bootstapping with 10,000 iterations. From Table 1, it is clear that we are able to obtain good test-retest reliability when calculating LIs in a typical manner. For all but the rhyme decision task, we see good test-retest reliability that exceeds our pre-registered thresold of *r*= .65. Furthermore, LIs for each task across the two sessions are lateralised at the participant level (all *p*s< .001). Specifically, the rhyme decision, dichotic listening, and finger tapping tasks were all lateralised to the left hemisphere (showing a left visual field advantage). The chimeric faces task was lateralised to the right hemipshere (showing a left visual field advantage). A highly similar pattern of results emerges for the *t*- and *z*-score measures, where each is lateralised at the population level (all *p*s< .001). The test-retest correlations are shown visually in Figure 5.

```{r make.combotable,echo=FALSE,include=TRUE}
#Make a table to summarise results for all 4 tasks
#This will gradually get populated as we go through each task
#Cells for each task are filled in course of plots.function chunk
combotable <- data.frame(matrix(NA,nrow=7,ncol=17))
colnames(combotable)<-c('.','RD 1','RD 2','DL 1','DL 2','CF 1','CF 2','FT 1','FT 2', 
                            'RD t 1','RD t 2','DL z 1','DL z 2','CF z 1','CF z 2','FT z 1','FT z 2')
combotable[,1]<-c('N','mean','SD','p-lateralised','p-normal','test-retest r','CI r')
```

```{r plots.function, echo=FALSE,include=TRUE}
# Adam's code here turned into generic function
# While we have the relevant columns specified here, I have added steps to populate combotable as we go.
# Also saving the scatterplots as png
densplots <- function(dat, time1,time2,name1,name2,scattername,combotable,combocolumn, y1, y2, x1, x2, ax, ay){

# plots
dens1<-ggplot(dat, aes(x=time1))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name1)
  dens1 <- dens1+ geom_vline(xintercept = 0)
# normality
shap1<-shapiro.test(time1)
ttest1 <- t.test(time1) #Check if significantly lateralised
Nt1 <- length(time1)-length(which(is.na(time1)))
combotable[1,combocolumn]<-Nt1
combotable[2,combocolumn]<-round(mean(time1,na.rm=T),2)
combotable[3,combocolumn]<-round(sd(time1,na.rm=T),2)
combotable[4,combocolumn]<-round(ttest1$p.value,3)
combotable[5,combocolumn]<-round(shap1$p.value,3)

# day 2
# plot denisty
dens2<-ggplot(dat, aes(x=time2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
  dens2 <- dens2+ geom_vline(xintercept = 0)
# normality
shap2<-shapiro.test(time2)
ttest2 <- t.test(time2)
Nt2 <- length(time2)-length(which(is.na(time2)))
myN<-min(Nt1,Nt2)
combotable[2,(combocolumn+1)]<-round(mean(time2,na.rm=T),2)
combotable[3,(combocolumn+1)]<-round(sd(time2,na.rm=T),2)
combotable[4,(combocolumn+1)]<-round(ttest2$p.value,3)
combotable[5,(combocolumn+1)]<-round(shap2$p.value,3)

# conduct Spearman's wth bootstrapped CI
myspear<-RVAideMemoire::spearman.ci(time1, time2, nrep = 10000, conf.level = 0.95)
  combotable[6,combocolumn]<-round(myspear$estimate,2)
  myci <- paste0(round(myspear$conf.int[1],2),' - ',round(myspear$conf.int[2],2))
  combotable[7,combocolumn]<-myci
# scatterplot time1 x time2
# position for label set to 50%x and 90%y
xpos <- max(time1,na.rm=T)*.5
ypos <- max(time2,na.rm=T)*.9

mycor<-cor.test(time1, time2, method = "spearman", alternative = "greater")$estimate

#Make a label to show the correlation with 95%CI and N on the plot
mycorlabel<-paste0('r = ',round(mycor,2),'\n(95% CI: ',round(myspear$conf.int[1],2),'-',round(myspear$conf.int[2],2),')\nN = ',myN)
#Find position in top L corner for legend
legposx <- .2 #not working - needs tweaking!
legposy <- .85
myscatter <- ggplot(dat, aes(x= time1, y= time2)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handed3)) + 
  geom_rug(aes(color= handed3)) +
  theme_classic() +
  ylab(name2) + 
  xlab(name1) +
  annotate("text", x = ax, y = ay, label = mycorlabel) +
  theme(legend.position = c(legposx,legposy), legend.title = element_blank()) + ylim(y1, y2) + xlim(x1, x2)
# Set relative size of marginal plots (main plot 10x bigger than marginals)
bigplot<-ggExtra::ggMarginal(myscatter, groupColour = TRUE, groupFill = TRUE)
scattername<-paste0(scattername,".png")
ggsave(scattername,plot=bigplot,dpi = 300, width = 6, height = 6,)

return(combotable)
}
```
<!--- We start by completing combotable for the 4 main measures. NB need to check polarity of tasks - also oddity re RDT distributions needs checking, and we said we'd do log transform , not sure we did--->
```{r do.plots,LI,echo=FALSE,include=TRUE}
# RHYME DECISION
# remove bad cases
RD.dat <- LIs[LIs$exRDT == 0,]
# now plot
dat <- RD.dat
time1<-RD.dat$Day1.RDT_RT_LI
time2<-RD.dat$Day2.RDT_RT_LI
name1<-"Rhyme decision LI (day 1)"
name2<-"Rhyme decision LI (day 2)"
scattername<-"RhymeDecisionScatter"
combocolumn<-2 #column to write day 1 summary for this task in combotable
y1= -12
y2= 12
x1= -12
x2= 12
ax= 8
ay= -8
combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)

# DICHOTIC LISTENING
# remove bad cases
DL.dat <- LIs[LIs$exclude == 0 | LIs$exclude == 1,]
# now plot
dat <- DL.dat
time1<-DL.dat$Day1_DL_acc_LI
time2<-DL.dat$Day2_DL_acc_LI
name1<-"Dichotic listening LI acc (day 1)"
name2<-"Dichotic listening LI acc (day 2)"
scattername<-"DichoticScatter"
combocolumn<-4 #column to write day 1 summary for this task in combotable
y1= -100
y2= 100
x1= -100
x2= 100
ax= 75
ay= -75
combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)

# CHIMERIC FACES
# remove bad cases
CF.dat <- LIs[LIs$exclude == 0 | LIs$exclude == 2,]
# now plot
dat <- CF.dat
time1<-CF.dat$Day1_CF_acc_LI
time2<-CF.dat$Day2_CF_acc_LI
name1<-"Chimeric faces LI acc (day 1)"
name2<-"Chimeric faces LI acc (day 2)"
scattername<-"ChimericScatter"
combocolumn<-6 #column to write day 1 summary for this task in combotabley1= -100
y1= -100
y2= 100
x1= -100
x2= 100
ax= 75
ay= -75
combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)

# FINGER TAPPING
dat <- LIs
time1<-LIs$Day1_finger_LI
time2<-LIs$Day2_finger_LI
name1<-"Finger tapping LI (day 1)"
name2<-"Finger tapping LI (day 2)"
scattername<-"FingerScatter"
combocolumn<-8 #column to write day 1 summary for this task in combotabley2= 100
y1= -35
y2= 35
x1= -35
x2= 35
ax= 26.25
ay= -26.25
combotable<-densplots(dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
```

```{r do.plot.t,echo=FALSE,include=TRUE}
# RHYME T-TEST
# now plot
dat <- RD.dat
time1<-RD.dat$logt1
time2<-RD.dat$logt2
name1<-"Rhyme decision t-score (day 1)"
name2<-"Rhyme decision t-score (day 2)"
scattername<-"rhyme.t"
combocolumn<-10 #column to write day 1 summary for this task in combotable
y1= -8
y2= 8
x1= -8
x2= 8
ax= 6
ay= -6
combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)

# DICHOTIC Z-TEST
# now plot
dat <- DL.dat
time1<-DL.dat$Day1_Dich_acc_Z
time2<-DL.dat$Day2_Dich_acc_Z
name1<-"Dichotic listening z-score (day 1)"
name2<-"Dichotic listening z-score (day 2)"
scattername<-"dichotic.z"
combocolumn<-12 #column to write day 1 summary for this task in combotable
y1= -20
y2= 20
x1= -20
x2= 20
ax= 15
ay= -15
combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)

# CHIMERIC FACES Z-TEST
# now plot
dat <- CF.dat
time1<-CF.dat$Day1_CF_acc_Z
time2<-CF.dat$Day2_CF_acc_Z
name1<-"Chimeric faces z-score (day 1)"
name2<-"Chimeric faces z-score (day 2)"
scattername<-"chimeric.z"
combocolumn<-14 #column to write day 1 summary for this task in combotable
y1= -20
y2= 20
x1= -20
x2= 20
ax= 15
ay= -15
combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)

# FINGER TAPPING Z-TEST
# now plot
dat <- LIs
time1<-LIs$Day1_Finger_acc_Z
time2<-LIs$Day2_Finger_acc_Z
name1<-"Finger tapping z-score (day 1)"
name2<-"Finger tapping z-score (day 2)"
scattername<-"finger.z"
combocolumn<-16 #column to write day 1 summary for this task in combotable
y1= -8
y2= 8
x1= -8
x2= 8
ax= 6
ay= -6
combotable<-densplots(LIs, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
```

Table 1. LIs and test-retest correlations for each taks.
```{r Table1,echo=FALSE}
#Needs more formatting. For word output may need to use flextable. See https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
kable(combotable) %>%
  kable_styling(c("bordered")) %>%
  add_header_above(c(" " = 1, "Typical LI" = 8, "t/z-score" = 8))
```
Note. LIs are transformed such that a negtaive value equates to a left lateralisation at the brain level. A positive value equates to right lateralisation at the brain level. RD= rhyme decision, DL= dichotic listening, CF: chimeric faces, and FT: finger tapping.

![Figure 5. Scatter plots for day 1 and day 2 data on each laterality task. The left column shows LIs calculated in the typical way. The right column shows LIs calculated using *t*- and *z*-scores. Here, distributions are shown for left and right handers, and an additional group of extreme left-handers who had a score below -90 on the Edinburgh Handedness Inventory.](combined_LI_cor.png)

Moving forward, we chose to use LIs calculated in a typical manner to address each remaining research question. To dtermine the smallest number of trials needed to achieve a test-retest reliabilty of *r*= .75, we recaclauted LIs by dropping one trial at a time for each participant...

We defined 'atypical laterality' as laterality that is in the opposite direction to the majority, and which has an absolute z-score or t-score value greater than 1.96. These z and t-scores indicate whether there is significant bias to the left or right within an individual. Based on our pre-registered analysis, we divided left-handers into two groups (1) those with a score of -90 or below on the Edinburgh Handedness Inventory (c.f. Mazoyer et al., 2014), (2) the remaining left handers. While our analyses inlcude only right handers and extreme left handers, we include the remaining left-handers in Table 2 for completeness.

```{r makecontingencies,echo=FALSE}
#We start by using z-scores to make a 3-way categorisation into:
#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
                       breaks = c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("TypLat", "NonLat", "AtypLat"), 
                       right = FALSE)
LIs$dichotic1.cut <- cut(LIs$Day1_Dich_acc_Z, 
                       breaks = c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("TypLat", "NonLat", "AtypLat"), 
                       right = FALSE)
LIs$chimeric1.cut <- cut(LIs$Day1_CF_acc_Z, 
                       breaks =  c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("AtypLat", "NonLat", "TypLat"), 
                       right = FALSE)
mytabr <- table(LIs$rhyme1.cut)
mytabd <- table(LIs$dichotic1.cut)
mytabc <- table(LIs$chimeric1.cut)
mytabc<-mytabc[c(3,2,1)]
mytab2 <-rbind(mytabr,mytabd,mytabc)
row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
mytab2<-data.frame(mytab2)
# combine tables for printing
mytab3 <- cbind(tabr, tabd, tabc)
# now do chi-squared tests
# rhyme
tabr <- table(LIs$handed3,LIs$rhyme1.cut)
rhy.chi <- chisq.test(tabr)
# dichotic
tabd <- table(LIs$handed3,LIs$dichotic1.cut)
dic.chi <- chisq.test(tabd)
# chimeric
tabc <- table(LIs$handed3,LIs$chimeric1.cut)
chim.chi <- chisq.test(tabc)
# BUT these don't actually run the planed tests. We run these by removing lefties here
#Planned test omitted group 'Left' and column Nonlat.
chi.r<-chisq.test(tabr[c(1,3),c(1,3)])
chi.d<-chisq.test(tabd[c(1,3),c(1,3)])
chi.c<-chisq.test(tabc[c(1,3),c(1,3)])
```

Table 2. Counts of laterality patterns for right-, left-, and extreme geft-handers acorss the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
```{r Table 2}
# plot group table
mytab3%>%
  kable() %>%  
  kable_styling(c("bordered")) %>%
  add_header_above(c(" " = 1, "Rhyme decision" = 3, "Dichotic listening" = 3, "Chimeric faces"= 3))
```
Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
